---
title: "exploration of dataset"
author: "Nick Lauerman"
date: "January 3, 2018"
output: 
  pdf_document: 
    fig_caption: yes
    fig_height: 7.5
    fig_width: 7.5
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# setup
```{r}
require(tidyverse)
```

# read data
```{r}
dataRaw <- read.csv("./data/online_sex_work.csv", 
                    stringsAsFactors = FALSE,
                    dec = ",")
dataRaw
```
# setting factors

Need to deal with NA prior to converting to factor


* Gender 
* Location 
* Verification
* Sexual_orientation 
* Sexual_polarity 
* Looking_for
* Risk 
```{r}
worker <- dataRaw
worker$Gender[is.na(worker$Gender)] <- "not_stated"
worker$Gender <- as.factor(worker$Gender)

worker$Location[is.na(worker$Location)] <- "unknown"
worker$Location <- as.factor(worker$Location)

worker$Verification[is.na(worker$Verification)] <- "unknown"
worker$Verification <- as.factor(worker$Verification)

worker$Sexual_orientation[is.na(worker$Sexual_orientation)] <- "not_stated"
worker$Sexual_orientation <- as.factor(worker$Sexual_orientation)

worker$Sexual_polarity[is.na(worker$Sexual_polarity)] <- "not_stated"
worker$Sexual_polarity <- as.factor(worker$Sexual_polarity)

worker$Looking_for[is.na(worker$Looking_for)] <- "unknown"
worker$Looking_for <- as.factor(worker$Looking_for)

worker$Risk[is.na(worker$Risk)] <- "unrated"
worker$Risk <- as.factor(worker$Risk)
```

# setting numeric
the following fields are characters and need to be converted to a number

* Points_Rank
* Number_of_Comments_in_public_forum
* Number_of_advertisments_posted
* Number_of_offline_meetings_attended
* Profile_pictures
* Age
* User_ID
```{r}
# worker$Points_Rank <- as.numeric(worker$Points_Rank) 
#      deal with value "a" "" and  1 196   1 215   1 437   1 626 see blanks in values

#worker$Number_of_Comments_in_public_forum <- as.numeric(worker$Number_of_Comments_in_public_forum)
#    deal with "" 1 174   1 275   1 894  2 062 

subset(worker, subset = Number_of_advertisments_posted == "")$Number_of_advertisments_posted <- "0" #explore gsub
worker$Number_of_advertisments_posted <- as.numeric(worker$Number_of_advertisments_posted)

subset(worker, Number_of_offline_meetings_attended == "")$NNumber_of_offline_meetings_attended <- "0"

worker$Number_of_offline_meetings_attended <- as.numeric(worker$Number_of_offline_meetings_attended)

subset(worker, Number_of_advertisments_posted == "")$Number_of_advertisments_posted <- "0"
worker$Profile_pictures <- as.numeric(worker$Profile_pictures)


#worker$Age <- as.numeric(worker$Age) 
#   Need to convert , to . as characters before converting to a number

#worker$User_ID <- as.numeric(worker$User_ID)
#   deal with ""
```

# Dates
comvert Member_since to a date
format is date.month(number).year(4 digits)
```{r}
require(lubridate)
worker$Member_since <- as.Date(worker$Member_since, format = "%d.%m.%Y")
```

