---
title: "exploration of dataset"
author: "Nick Lauerman"
date: "January 3, 2018"
output: 
  pdf_document: 
    fig_caption: yes
    fig_height: 7.5
    fig_width: 7.5
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# setup
```{r required libraries}
require(tidyverse)
require(lubridate)
```

# read data
```{r read data, cache=TRUE,cache.path="./cache/"}
dataRaw <- read.csv("./data/online_sex_work.csv", 
                    stringsAsFactors = FALSE)
#dataRaw
```

# Remove blank rows
the data set as published has many rows at the end that are full of blank "" 
values. These rows are being deleted based on the User ID value being selected as 
not being "".
```{r reset point}
worker <- dataRaw
```
```{r remove blank rows}
worker <- subset(worker, subset = User_ID != "")
```



# setting factors

Need to deal with NA prior to converting to factor


* Gender 
* Location 
* Verification
* Sexual_orientation 
* Sexual_polarity 
* Looking_for
* Risk 
```{r remove NA for factors}

worker$Gender[is.na(worker$Gender)] <- "not_stated"
worker$Gender <- as.factor(worker$Gender)

#need to deal with blank values
worker$Location[is.na(worker$Location)] <- "unknown"
worker$Location <- as.factor(worker$Location)

worker$Verification[is.na(worker$Verification)] <- "unknown"
worker$Verification <- as.factor(worker$Verification)

worker$Sexual_orientation[is.na(worker$Sexual_orientation)] <- "not_stated"
worker$Sexual_orientation <- as.factor(worker$Sexual_orientation)

worker$Sexual_polarity[is.na(worker$Sexual_polarity)] <- "not_stated"
worker$Sexual_polarity <- as.factor(worker$Sexual_polarity)

#need to deal with blank values
worker$Looking_for[is.na(worker$Looking_for)] <- "unknown"
worker$Looking_for <- as.factor(worker$Looking_for)

worker$Risk[is.na(worker$Risk)] <- "unrated"
worker$Risk <- as.factor(worker$Risk)
```



# setting numeric
the following fields are characters and need to be converted to a number

* Number_of_advertisments_posted
* Number_of_offline_meetings_attended
* Profile_pictures
* User_ID
* Age
* Number_of_Comments_in_public_forum
* Points_Rank
```{r convert to numeric values}

worker$Number_of_advertisments_posted <- as.integer(worker$Number_of_advertisments_posted)

worker$Number_of_offline_meetings_attended <- as.integer(worker$Number_of_offline_meetings_attended)

worker$Profile_pictures <- as.integer(worker$Profile_pictures)

worker$User_ID <- as.integer(worker$User_ID)

worker$Age <- gsub(pattern = ",",
                   replacement = ".",
                   worker$Age)
worker$Age <- as.numeric(worker$Age)
# NA equate to unknown

worker$Number_of_Comments_in_public_forum <- gsub(pattern = " ",
                                             replacement = "",
                                             worker$Number_of_Comments_in_public_forum)
worker$Number_of_Comments_in_public_forum <- as.numeric(worker$Number_of_Comments_in_public_forum)

worker$Points_Rank <- gsub(pattern = "a",
                      replacement = "0",
                      worker$Points_Rank)
worker$Points_Rank <- gsub(pattern = " ",
                      replacement = "",
                      worker$Points_Rank)
worker$Points_Rank <- as.integer(worker$Points_Rank)
```

# Dates
comvert Member_since to a date
format is date.month(number).year(4 digits)
```{r}
require(lubridate)
worker$Member_since <- as.Date(worker$Member_since, format = "%d.%m.%Y")
```

